Lake Photic Zone Temperature across the Conterminous United States

### 2020-08-07 manuscript/manuscript.Rmd
new figure

---
title: "Lake Photic Zone Temperature across the Conterminous United States"
author:
- Kreakie, B. J. ^\*^ ^*1*^,  Shivers, S. ^*2*^, Hollister. J. W. ^*1*^, Milstead,
  W. Bryan. ^*1*^,
- ^*1*^ US Environmental Protection Agency, Office of Research and Development,  Atlantic
  Coastal Environmental Sciences Division (ACESD), Narragansett, RI 02882
- ^*2*^ ORISE, Narragansett, RI 02882
- '^\*^  *corresponding author: kreakie.betty@epa.gov*'
bibliography: manuscript.bib
output:
  word_document:
    reference_docx: manuscript_template.docx
  pdf_document: default
editor_options:
  chunk_output_type: inline
csl: plos.csl
abstract: As the average global air temperature on Earth increases, surface temperatures
  of lakes are also increasing globally (0.34 °C per decade from 1985 to 2009).  The
  influence of this increased temperature touches all biotic and abiotic components
  of lentic ecosystems. Here we present a simple yet robust model of lake photic zone
  temperature using the 2007 and 2012 US EPA’s National Lakes Assessment data for
  the conterminous United States. The final model has a mean square error of 2.19
  and an adjusted R(n<sup>2</sup>) of 0.88. The sampling date, that day’s average ambient air temperature,
  data obtained from the PRISM Climate Group, and longitude are the most important
  variables impacting the final model’s accuracy. Given the importance of temperature
  to a lake ecosystem, especially to cyanobacteria bloom dynamics, this model can
  be a valuable tool for researchers and lake resource managers. Daily predicted lake
  photic zone temperature for all lakes in the conterminous US can now be straightforwardly
  estimated based on basic ambient temperature and location information.
---

@@

Figure 1: Map of 2007 and 2012 US EPA’s National Lake Assessment (NLA) Lakes

>>>>

---
title: "Lake Photic Zone Temperature across the Conterminous United States"
author:
- Kreakie, B. J. ^\*^ ^*1*^,  Shivers, S. ^*2*^, Hollister. J. W. ^*1*^, Milstead,
  W. Bryan. ^*1*^,
- ^*1*^ US Environmental Protection Agency, Office of Research and Development,  Atlantic
  Coastal Environmental Sciences Division (ACESD), Narragansett, RI 02882
- ^*2*^ ORISE, Narragansett, RI 02882
- '^\*^  *corresponding author: kreakie.betty@epa.gov*'
bibliography: manuscript.bib
output:
  pdf_document: default
  word_document:
    reference_docx: manuscript_template.docx
editor_options:
  chunk_output_type: inline
csl: plos.csl
abstract: As the average global air temperature on Earth increases, surface temperatures
  of lakes are also increasing globally (0.34 °C per decade from 1985 to 2009).  The
  influence of this increased temperature touches all biotic and abiotic components
  of lentic ecosystems. Here we present a simple yet robust model of lake photic zone
  temperature using the 2007 and 2012 US EPA’s National Lakes Assessment data for
  the conterminous United States. The final model has a mean square error of 2.19
  and an adjusted R(n<sup>2</sup>) of 0.88. The sampling date, that day’s average ambient air temperature,
  data obtained from the PRISM Climate Group, and longitude are the most important
  variables impacting the final model’s accuracy. Given the importance of temperature
  to a lake ecosystem, especially to cyanobacteria bloom dynamics, this model can
  be a valuable tool for researchers and lake resource managers. Daily predicted lake
  photic zone temperature for all lakes in the conterminous US can now be straightforwardly
  estimated based on basic ambient temperature and location information.
---

@@

![`r fig_num("LakePhoticZonev2")`](../figures/LakePhoticZonev2.jpg)
Figure 1: Map of 2007 and 2012 US EPA’s National Lake Assessment (NLA) Lakes
